---
title: "test-did"
author: "Dan Ovando"
date: "12/5/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rstanarm)
```

The idea here is to simulate the concept behind the ahnold regression before you go crazy coding new things to make sure that it actually works. So, your goal is to simulate two populations with a shared trend, one of which gets a treatment that phases in/shocks in, and use DiD to estimate the effect


```{r}



common_noise <- data_frame(year = 0:10) %>% 
  mutate(enso = sin(year) * 50)

common_slope = 1.1

mpa_effect <- 10

mpa_slope <- 20

mean_noise <- 20

mpa_year <- 5

sigma_pop <- 5

n_species <- 30

life_history <-
  data_frame(
  species = 1:n_species,
  slope_effect = rlnorm(n_species),
  enso_effect = sample(c(-1, 1), n_species, replace = T),
  targeted = sample(c(0, 1), n_species, replace = T),
  shared_slope = common_slope,
  intercept = runif(n_species, 10,200) + 75*targeted
  )
  
populations <- cross_df(list(species = 1:n_species, year = 0:10)) %>% 
  left_join(life_history, by = 'species') %>% 
  left_join(common_noise, by = 'year') 


populations <- populations %>% 
  mutate(population = intercept + year * (shared_slope * slope_effect) + (targeted * mpa_effect) * (year >= mpa_year) + (targeted * mpa_slope) * pmax(0,year - mpa_year) + rnorm(nrow(.),0,sigma_pop) + enso * enso_effect) %>% 
  mutate(targeted = targeted == 1,
         post_mpa = year >= mpa_year,
         factor_year = as.factor(year),
         mpa_bump = mpa_effect * (year >= mpa_year),
         mpa_trend = mpa_slope * pmax(0,year - mpa_year),
         mpa_net = mpa_bump + mpa_trend,
         numeric_targeted = as.numeric(targeted),
         species = as.factor(species))

populations %>% 
  ggplot(aes(year,population, color = targeted)) + 
  geom_line(aes(group = interaction(species,targeted))) + 
  geom_point() +
  geom_vline(aes(xintercept = mpa_year))

```

Now, convince yourself that the classic DiD can do this... yes it can hooray

```{r}
simple_did <- lm(population ~ targeted + post_mpa + targeted:post_mpa, data = populations)

summary(simple_did)

```


```{r}

simple_stan_did <- rstanarm::stan_glm(population ~ targeted + post_mpa + targeted:post_mpa, data = populations, chains = 1)

summary(simple_stan_did)

```


Now, let's try it with the time effects

```{r}

time_did <- lm(population ~ factor_year + targeted + targeted:factor_year, data = populations)

summary(time_did)

broom::tidy(time_did) %>%
  filter(str_detect(term, 'targetedTRUE')) %>%
  mutate(year = map_dbl(term, ~str_replace_all(.x,'\\D','') %>% as.numeric())) %>%
  filter(!is.na(year)) %>% 
  ggplot(aes(year, estimate)) + 
  geom_point()

populations %>% 
  filter(targeted == T) %>% 
  ggplot(aes(year, mpa_net)) + 
  geom_point()
```



```{r wi-enso}


time_did <- rstanarm::stan_glm(population ~ factor_year + numeric_targeted + numeric_targeted:factor_year + enso:species, data = populations, chains = 1)

summary(time_did)

did_terms <- time_did$stan_summary %>%
  as.data.frame() %>%
  mutate(term = rownames(.)) %>%
  filter(str_detect(term, 'numeric_targeted')) %>%
  mutate(year = map_dbl(term, ~str_replace_all(.x,'\\D','') %>% as.numeric())) %>%
  filter(!is.na(year))

did_terms %>%
  ggplot() +
  geom_hline(aes(yintercept = 0), color = 'red') +
  geom_vline(aes(xintercept = mpa_year), color = 'blue') +
  geom_pointrange(aes(
    x = year,
    y = mean,
    ymin = `2.5%`,
    ymax = `97.5%`
  ))


```


```{r wo-enso}


time_did <- rstanarm::stan_glm(population ~ factor_year + numeric_targeted + numeric_targeted:factor_year, data = populations, chains = 1)

summary(time_did)

did_terms <- time_did$stan_summary %>%
  as.data.frame() %>%
  mutate(term = rownames(.)) %>%
  filter(str_detect(term, 'numeric_targeted')) %>%
  mutate(year = map_dbl(term, ~str_replace_all(.x,'\\D','') %>% as.numeric())) %>%
  filter(!is.na(year))

did_terms %>%
  ggplot() +
  geom_hline(aes(yintercept = 0), color = 'red') +
  geom_vline(aes(xintercept = mpa_year), color = 'blue') +
  geom_pointrange(aes(
    x = year,
    y = mean,
    ymin = `2.5%`,
    ymax = `97.5%`
  ))


```

```{r wi-enso-species}


time_did <- rstanarm::stan_glmer(population ~ factor_year + numeric_targeted + numeric_targeted:factor_year + (1 + enso|species), data = populations, chains = 1)

summary(time_did)

did_terms <- time_did$stan_summary %>%
  as.data.frame() %>%
  mutate(term = rownames(.)) %>%
  filter(str_detect(term, 'numeric_targeted')) %>%
  mutate(year = map_dbl(term, ~str_replace_all(.x,'\\D','') %>% as.numeric())) %>%
  filter(!is.na(year))

did_terms %>%
  ggplot() +
  geom_hline(aes(yintercept = 0), color = 'red') +
  geom_vline(aes(xintercept = mpa_year), color = 'blue') +
  geom_pointrange(aes(
    x = year,
    y = mean,
    ymin = `2.5%`,
    ymax = `97.5%`
  ))


```



# Generation DiD

Cool, now try it again with generational DiD and make sure that it works like you think it does

```{r}



common_noise <- data_frame(year = 0:10) %>% 
  mutate(enso = sin(year) * 50)

common_slope = 1.1

mpa_effect <- 10

mpa_slope <- 20

mean_noise <- 20

mpa_year <- 5

sigma_pop <-10

n_species <- 30

life_history <-
  data_frame(
  species = 1:n_species,
  slope_effect = rlnorm(n_species),
  enso_effect = sample(c(-1, 1), n_species, replace = T),
  targeted = sample(c(0, 1), n_species, replace = T),
  shared_slope = common_slope,
  intercept = runif(n_species, 100,400) + 75*targeted,
    age_mature = runif(n_species,1,1)

  # age_mature = runif(n_species,1,3 + 2*targeted)
  )
  
populations <- cross_df(list(species = 1:n_species, year = 0:10)) %>% 
  left_join(life_history, by = 'species') %>% 
  left_join(common_noise, by = 'year') 


populations <- populations %>% 
  mutate(population = intercept + year * (shared_slope * slope_effect) + (targeted * mpa_effect) * (year >= mpa_year) + (targeted * mpa_slope) * pmax(0,round((year - mpa_year) / age_mature)) + rnorm(nrow(.),0,sigma_pop) + enso * enso_effect) %>% 
  mutate(generations_protected = pmax(0,round((year - mpa_year)/age_mature)),
    targeted = targeted == 1,
         post_mpa = year >= mpa_year,
         factor_year = as.factor(year),
         mpa_bump = mpa_effect * (year >= mpa_year),
         mpa_trend = mpa_slope * generations_protected,
         mpa_net = mpa_bump + mpa_trend,
         numeric_targeted = as.numeric(targeted),
         species = as.factor(species)) %>% 
  mutate(generations_protected = as.factor(generations_protected))

populations %>% 
  ggplot(aes(year,population, color = targeted)) + 
  geom_line(aes(group = interaction(species,targeted))) + 
  geom_point() +
  geom_vline(aes(xintercept = mpa_year))

populations %>% 
  ggplot(aes(year,mpa_net)) + 
  geom_point()

populations %>% 
  ggplot(aes(generations_protected,mpa_net)) + 
  geom_point()

```


Now, try sampling from that population. The idea here is to generate a bunch of sampling events. At each event, you have some probability of detecting the species, based on environmental covariates and the relative abundance of that species. If you detect it, you draw an estimate of the populatoin in that year with mean true and sigma something. 

That will leave you with a database of observations, that you can then plug into your delta GLM approach to see if it actually works, throwing in a covariate that helps with detection, aka visibility, that is unrelated to other things. So, in theory then the probability of detection should reflect abundance resulting from the MPA as well as the other covariate. 

```{r}


go_sample <- function(pops, events){
  
sampling_events <- data_frame(event = 1:events, vis = rnorm(events, 0,4)) %>% 
  mutate(vis = center_scale(vis))

sample_foo <- function(vis, pops, vis_beta = 1, pop_beta = 1, obs_cv = .25){
  pops <- pops %>% 
    ungroup() %>% 
    mutate(prob_seen = 1 / (1 + exp(-(0 + vis * vis_beta + pop_beta * cs_population)))) %>% 
    mutate(seen = rbinom(nrow(pops), 1, prob_seen)) %>% 
    mutate(pop_hat = seen * rnorm(nrow(pops), population, obs_cv * population))
           
           # sqrt(log(obs_cv^2 +1)))
}

sampling_events <- sampling_events %>% 
  mutate(samples = map(vis, sample_foo, pops = pops)) %>% 
  unnest()
  
} # close go sample

events <- 200

set.seed(42)

samples <- populations %>% 
  mutate(cs_population = center_scale(population)) %>% 
  nest(-year) %>% 
  mutate(samples = map(data, go_sample, events = 200)) %>% 
  select(year,samples) %>% 
  unnest()


 samples %>% 
   ggplot() +
   geom_smooth(aes(year, pop_hat, color = species))
 
 
 samples %>% 
   ggplot(aes(cs_population, prob_seen)) + 
   geom_point()
 
  samples %>% 
   ggplot(aes(vis, prob_seen)) + 
   geom_point()



```


OK! That all works nicely... now the annoying part is that you'll need to pass this all to TMB somehow to to fit the model really... ugh. 


```{r}
data <- samples %>% 
  mutate(any_seen = pop_hat > 0) %>% 
  mutate(classcode = species) %>% 
  mutate(log_density  = (population))

non_nested_variables <- c(
  'vis',
  'factor_year',
  'targeted'
)

seen_has_important <- data %>%
  filter(any_seen == T) %>%
  select(non_nested_variables) %>%
  mutate(index = 1:nrow(.)) %>%
  na.omit()

seeing_has_important <- data %>%
  select(non_nested_variables) %>%
  mutate(index = 1:nrow(.)) %>%
  na.omit()

numeric_species_key <-
  data_frame(classcode = unique(data$classcode)) %>%
  arrange(classcode) %>%
  mutate(numeric_classcode = 1:nrow(.))

seen_data <- data %>%
  filter(any_seen == T) %>%
  left_join(numeric_species_key, by = "classcode") %>%
  slice(seen_has_important$index)

log_density <- (seen_data$population)


seeing_data <- data %>%
  left_join(numeric_species_key, by = "classcode") %>%
  slice(seeing_has_important$index)

any_seen <- seeing_data$any_seen

x_seen_non_nested <- seen_data %>%
  select(non_nested_variables) %>%
  mutate(intercept = 1) %>%
  demons::spread_factors(drop_one = T) %>%
  purrrlyr::dmap(center_scale) %>%
  mutate(targeted = seen_data$targeted)

x_seen_did <- seen_data %>%
  select(targeted, factor_year) %>%
  mutate(index = 1:nrow(.)) %>%
  spread(factor_year, targeted, fill = 0) %>%
  arrange(index) %>%
  select(-(1:2)) # drop index and one factor level

x_seen <- bind_cols(x_seen_non_nested, x_seen_did) %>% 
  mutate(log_density = log_density, 
         targeted = as.numeric(targeted)) 

check_seen <- lm(log_density ~ . - 1, data = x_seen)


# prepare seeing ----------------------------------------------------------

x_seeing_non_nested <- seeing_data %>%
  select(non_nested_variables) %>%
  mutate(intercept = 1) %>%
  demons::spread_factors(drop_one = T) %>%
  purrrlyr::dmap(center_scale) 

x_seeing_did <- seeing_data %>%
  select(targeted, factor_year) %>%
  mutate(index = 1:nrow(.)) %>%
  spread(factor_year, targeted, fill = 0) %>%
  arrange(index) %>%
  select(-(1:2)) # drop index and one factor level

x_seeing <- bind_cols(x_seeing_non_nested, x_seeing_did) %>% 
  mutate(any_seen = any_seen %>% as.numeric(), 
         targeted = as.numeric(targeted)) 

check_seeing <- glm(any_seen ~ . - 1, data = x_seeing, family = binomial)

# create standard matrix --------------------------------------------------

standard_non_nested <- x_seeing_non_nested %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(mean_value = mean(value, na.rm = T)) %>%
  spread(variable, mean_value)

standard_non_nested <-
  map_df(seq_along(unique(seeing_data$factor_year)), function(x, sdata) {
    sdata
  }, sdata = standard_non_nested) %>%
  ungroup() %>%
  select(colnames(x_seeing_non_nested)) %>%
  slice(-1) # make sure things are in the same column order as they used to be

standard_did <- data_frame(year = unique(seeing_data$factor_year)) %>%
  spread_factors(drop_one = T) %>%
  slice(-1)

standard <- bind_cols(standard_non_nested, standard_did)


# check_did

did_seen <- check_seen %>% broom::tidy() %>% 
  mutate(term = str_replace_all(term, "`", "")) %>% 
  filter(str_detect(term,"\\d") & !str_detect(term,"factor_year"))
 

did_seeing <- check_seeing %>% broom::tidy()

prob_seeing <- 1 / (1 + exp(-(as.matrix(standard) %*% did_seeing$estimate)))


plot(did_seen$estimate)

plot(prob_seeing)

plot(did_seen$estimate * prob_seeing)


```



```{r simple-generations}


time_did <- rstanarm::stan_glm(population ~ numeric_targeted + generations_protected + factor_year + numeric_targeted:generations_protected, data = populations, chains = 1)

summary(time_did)

did_terms <- time_did$stan_summary %>%
  as.data.frame() %>%
  mutate(term = rownames(.)) %>%
  filter(str_detect(term, 'numeric_targeted')) %>%
  mutate(year = map_dbl(term, ~str_replace_all(.x,'\\D','') %>% as.numeric())) %>%
  filter(!is.na(year))

did_terms %>%
  ggplot() +
  geom_hline(aes(yintercept = 0), color = 'red') +
  geom_vline(aes(xintercept = mpa_year), color = 'blue') +
  geom_pointrange(aes(
    x = year,
    y = mean,
    ymin = `2.5%`,
    ymax = `97.5%`
  ))


```

```{r wi-enso-species}


time_did <- rstanarm::stan_glmer(population ~ factor_year +generations_protected + numeric_targeted + numeric_targeted:generations_protected + (1 + enso|species), data = populations, chains = 1)

summary(time_did)

did_terms <- time_did$stan_summary %>%
  as.data.frame() %>%
  mutate(term = rownames(.)) %>%
  filter(str_detect(term, 'numeric_targeted')) %>%
  mutate(year = map_dbl(term, ~str_replace_all(.x,'\\D','') %>% as.numeric())) %>%
  filter(!is.na(year))

did_terms %>%
  ggplot() +
  geom_hline(aes(yintercept = 0), color = 'red') +
  geom_vline(aes(xintercept = mpa_year), color = 'blue') +
  geom_pointrange(aes(
    x = year,
    y = mean,
    ymin = `2.5%`,
    ymax = `97.5%`
  ))



populations %>% 
  ggplot(aes(generations_protected,mpa_net)) + 
  geom_point()

```






```{r}

```

