---
title: Regional Effects of Marine Protected Areas May Be Small and Difficult to Detect

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Daniel Ovando
    affiliation: a,1,2
  - name: Bob Security
    affiliation: a,b
address:
  - code: a
    address: University of Washington, School of Aquatic and Fishery Science
  - code: b
    address: Another University Department, Street, City, State, Zip

corresponding_author:
  - code: 2
    text: "To whom correspondence should be addressed. E-mail: danovan@uw.edu"

# For footer text
lead_author_surname: Ovando


author_contributions: |
  Please provide details of author contributions here.

## Remove this if not required
conflict_of_interest: |
  Please declare any conflict of interest here.

abstract: |
  Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.

  Make the case clearer why this is important to know, can't design and monitor what you don't understand

  Marine Protected Areas (MPAs) cover XX% of the worlds oceans, a XX% increase in area over the last XX years. The XX goals call for a further increase to XX% by XX. This expansion is driven by a desire to achieve conservation benefits inside and outside of MPA borders, as well as economic benefits through tourism and fisheries spillover. While the conservation effects of MPAs inside their borders are well demonstrated, the effects of MPAs on the broader ecosystems and fisheries around them are far less understood. We utilize theory and data to ask, what should we expect the regional effects of MPAs to be, and how can we detect them? We integrated multiple theories of MPA effects into a cohesive bio-economic simulator that allows us to explore the regional conservation and fishery effects of MPAs across range of realistic fishery scenarios. The mean effect of MPAs on total population sizes (inside and outside of MPAs) was a XX% increase in biomass, relative to a world in which the MPAs did not exist (equivalent to recovering xx% of unfished biomass). We demonstrate the challenges of estimating an effect of this size in natural systems using 15xx years of data on the density of fishes inside and outside of MPAs within the Channel Islands National Marine Sanctuary. We find no detectable effect of MPAs throughout the study region, though we can rule out conservation benefits above XX (or losses below XX). Our results show that the regional effects of MPAs may be much smaller than previously thought, requiring us to think carefully about their role in marine resource management.

significance: |


  Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.

acknowledgements: |
  Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.

keywords:
  - Marine Protected Areas
  - Conservation
  - Fisheries
  - optional
  - optional

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: pnas-sample.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: false

watermark: false

output: rticles::pnas_article
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "CairoPDF", dpi = 600)
# knitr::opts_knit$set(root.dir = "/Users/danovando/PhD/zissou") # not entirely reprodcible hack to deal with knit child in dissertation.. not idealk

```


```{r, include=FALSE}
library(sf)
library(ggmap)
library(viridis)
library(hrbrthemes)
library(scales)
library(patchwork)
library(rpart)
library(extrafont)
library(caret)
library(ggsci)
library(rEDM)
library(tidyverse)
extrafont::loadfonts()
functions <- list.files(here::here("functions"))

walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
save_plots <- TRUE

sim_years <- 50

burn_years <- 20

num_patches <- 25

  samps <- 20000
 
  run_name <- "v3.0"

run_dir <- here::here("results", run_name)

experiment_dir <- here::here("results",run_name, "experiments")

load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))

load(file = here::here("results",run_name, "model_runs.Rdata"))

load(file = here::here("results",run_name,"processed_grid.Rdata"))

load(file = here::here("results",run_name,"about_time.Rdata"))

channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))

ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>% 
  rmapshaper::ms_simplify() %>% 
  sf::st_transform(crs = 4326)


  zissou_theme <- theme_ipsum(base_size = 12, axis_title_size = 12, strip_text_size = 12)

  theme_set(zissou_theme)
  

models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)

model_runs <- model_runs %>%
  filter(models_worked) %>%
  mutate(tmb_fit = map(tmb_fit,"result")) %>%
  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
  mutate(did_plot = map(processed_fits, "did_plot"))

base_run <- model_runs %>% 
filter(var_names == "pisco_a", mpa_only == FALSE, center_scale == TRUE)

mpa_run <- model_runs %>% 
  filter(data_source == "pisco",
         var_names == "pisco_a",
         mpa_only == TRUE)

kfm_run <- model_runs %>% 
  filter(data_source == "kfm")
  
fitted_data <- base_run$data[[1]]

zissou_fit <- base_run$tmb_fit[[1]]

report <- base_run$tmb_fit[[1]]

site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
  magrittr::set_colnames(., tolower(colnames(.))) %>%
  select(
    site,
    side,
    mpagroup,
    mpa_status,
    reserve,
    region,
    year_mpa,
    mpaareanm2,
    lat_wgs84,
    lon_wgs84
  ) %>%
  rename(lat_wgs84 = lon_wgs84,
         lon_wgs84 = lat_wgs84) %>%
  unique() %>%
  mutate(eventual_mpa = (year_mpa > 0))


 seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
   filter(variable == "seen_non_nested_betas") %>%
   rename(group = variable) %>%
   mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested%>% colnames())

 seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
   filter(variable == "seeing_non_nested_betas") %>%
   rename(group = variable) %>%
   mutate(variable  = zissou_fit$zissou_data$x_seeing_non_nested %>% colnames())

 seen_year_species_betas <- zissou_fit$zissou_estimates %>%
   filter(variable == "seen_year_species_betas") %>%
   rename(group = variable) %>%
   mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

 seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
   filter(variable == "seeing_year_species_betas") %>%
   rename(group = variable) %>%
   mutate(variable =zissou_fit$zissou_data$x_seeing_year_species %>% colnames())

 seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
   filter(variable == "seen_region_cluster_betas") %>%
   rename(group = variable) %>%
   mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

 seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
   filter(variable == "seeing_region_cluster_betas") %>%
   rename(group = variable) %>%
   mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())


 did_betas <- zissou_fit$zissou_estimates %>%
   filter(variable == "mpa_effect") %>%
   mutate(group = variable) %>%
   mutate(year = fitted_data$year %>% unique())


 betas <- bind_rows(
   seen_non_nested_betas,
   seeing_non_nested_betas,
   seen_year_species_betas,
   seeing_year_species_betas,
   seen_region_cluster_betas,
   seeing_region_cluster_betas,
   did_betas %>% select(-year)
 ) %>%
   as_data_frame()

 non_nested_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested")) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x = variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_wrap(~group) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10))


 year_species_effects_plots <- betas %>%
   filter(str_detect(group, "year_species_betas")) %>%
   mutate(year = str_replace_all(variable,"\\D","") %>% as.numeric()) %>%
   mutate(classcode = str_split(variable,'-', simplify = T)[,2]) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = classcode), alpha = 0.25) +
   geom_line(aes(x = year, y = estimate, color = classcode)) +
   facet_wrap(~group)

  region_cluster_plots <- betas %>%
   filter(str_detect(group, "region_cluster_betas")) %>%
   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
   ggplot() +
    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
    geom_pointrange(aes(x = region,
                        y = estimate,
                        ymin = lower,
                        ymax = upper, color = cluster)) +
    facet_wrap(~group)



 did_plot <- did_betas %>%
   ggplot() +
   geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
 geom_hline(aes(yintercept = 0)) +
   geom_pointrange(aes(
     year,
     y = estimate,
     ymin = lower,
     ymax = upper
   ),
   size = 1.5) +
   labs(x = "Year", y = "Targeted Trend Divergence")

top_species <- base_run$data[[1]]$classcode %>% unique()

cip_data <- pisco_data %>% 
  left_join(site_data, by = c("site","side")) %>% 
  filter(is.na(eventual_mpa) == F) %>% 
  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
         classcode %in% top_species)

# process outcomes

outcomes <- processed_grid %>%
  select(-fishery_effect) %>% 
  slice(1:samps) %>% 
  unnest() %>% 
  mutate(year = year - burn_years) %>% 
  mutate(years_protected = year - year_mpa + 1) %>% 
  mutate(mpa_effect = pmin(mpa_effect,2)) %>% 
  group_by(experiment) %>% 
  mutate(b0 = `no-mpa`[year == min(year)]) %>% 
  mutate(depletion = 1 - `no-mpa`/b0) %>% 
  ungroup() %>% 
  mutate(pop_effect = pmin(1,(`with-mpa` - `no-mpa`) / b0))

eqo <- outcomes %>% 
  filter(year == max(year))

fishery_outcomes <- processed_grid %>%
  select(-mpa_effect) %>% 
  slice(1:samps) %>% 
  unnest() %>% 
  mutate(year = year - burn_years) %>% 
  mutate(years_protected = year - year_mpa + 1) %>% 
  mutate(fishery_effect = pmin(mpa_effect,1)) %>% 
  left_join(outcomes %>% select(experiment, year, depletion), by = c("experiment", "year")) %>% 
  filter(fleet_model != "constant-catch") %>% 
  mutate(msy_effect = pmin(1,(`with-mpa` - `no-mpa`) / msy))

fishery_eqo <- fishery_outcomes %>% 
  filter(year == max(year))


facet_labels <- c(
  mpa_size = "Range in MPA",
  depletion = "Pre-MPA Depletion"
)

```


xx mention Ferraro et al

No-take Marine Protected Areas (MPAs, which we will define here as spatial regions in the ocean in which fishing for species of interest is prohibited) have a long history in the management of marine resources. Traditional cultures in Oceania utilized (often temporary) MPAs as a sort of "fish bank" for times of need [@Johannes1978]. Modern MPAs were first put in place primarily as conservation areas, analogs to terrestrial reserves deigned to protect iconic landscapes such as Yellowstone or Kruger National Parks [citation]. However, over time our goals and expectations for MPAs have evolved; MPAs are now frequently considered to both protect the marine ecosystems within their boundaries, and bolster fish populations and fishing opportunities in their surrounding waters [@Gaines2010].  


With these goals in mind, the XX has called for a an XX% increase in the XX of global MPAs by XX [citation]. Our understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited though. We have clear evidence that well enforced and appropriately sized MPAs can provide conservation benefits within their borders [@Lester2009;@Halpern2003; @Edgar2014; @Gerber2005]. As these conservation benefits accrue inside one or a network of MPAs, the MPA(s) can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas. Several studies have documented empirical evidence for the existence of adult or larval spillover affecting both abundance and fisheries [e.g. @Goni2010; @Halpern2009;@Kay2012; @Stobart2009; @McClanahan2000; @Russ1996; @Thompson2017]. The more complex and potentially important question is not whether spillover occurs, but what the net effect of spillover is. From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds? From a conservation perspective, does the buildup of adults inside an MPA increase abundance outside through spillover, or does concentration of fishing outside the reserve result in a net loss in regional abundance?

As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, it is critically important that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects. To address this gap, this study examines two critical questions: 1) What do we expect the regional-scale conservation and fishery effects of MPAs to be and 2) When (and how) can we expect to detect these effects? We address these questions using a simulation analysis to frame the theoretical regional conservation and fishery impacts of MPAs, from which we then develop an empirical assessment of the evidence for regional-level conservation effects of MPAs resulting from a network of closures put in place in the Channel Islands, California, in 2003. 

Why should we care about MPAs?

what is currently known?

theory and evidence

what do we contribute

novelty of simulation method

identifications strategy



What Should We Expect From MPAs? {.unnumbered}
----------------------

Much of the literature on the effects of Marine Protected Areas focsuses on assessing the conservation effects within the borders of protected aras [e.g. @Edgar2014]. While these effects are important, no MPA is truly an island; rather the organisms within the borders of protected areas are generally part of a broader population, the biological stock, connected through adult movement and larval dispersal. With this in mind then, a broader question we should ask of MPAs is not just are there more fish inside the borders of a reserve, but how do MPAs affect the broader region in which they sit? 

For the cases of conservation outcomes, we define these regional effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (other metrics such as biodiversity are also important but are beyond the scope of this analysis). For the fisheries case we consider the total fishery catches with the MPAs realtive to the fishery catches that would have occured without the MPAs. Note that these definitions are different...

XX link to ferraro, explain why different than the CPUE spillover papers

XX explain the example MPA scenario


Regional Conservation Effects {#conservation .unnumbered}
-------------------

What does basic theory and intuition suggest should be the magnitude of these regional effects? On one hand, if we imagine a region that has driven its fish populations to near extinction, that then places 100% of its waters inside a no-take MPA, we would expect the regional-scale conservation effects to be massive. On the other hand, if we implement an MPA for a lightly fished sedentary species it is possible to create a net conservation loss as fishing pressure is concentrated outside the reserve. Within these broad bounds, numerous factors can act to affect the regional effects of MPAs. These include the scale of adult and larval dispersal relative to the size of the MPAs [@Gaines2003; @Botsford2008; @DiFranco2018], the strength and timing of density dependence in the population (e.g. pre or post settlement), how overfished the population was pre-MPA, and how fishing activity responds to the implementation of the MPAs [@Hilborn2004; @Hilborn1992; @Walters2004; @Hastings2003;@Gerber2003; @Gerber2005; @Hilborn2004a; @Gaines2010]. In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@Gaines2010; @Costello2010]. Broadly, a wide range of theory and modeling exercises indicate that the expected regional conservation effects of MPAs can vary widely and are extremely context- dependent [@Fulton2015]. 

These prior works have either focused on the effect of a select set of drivers (such as dispersal rates) on conservation outcmoes, or have been highly locallized assessments that incorporate multiple drivers but are tuned to one specific region. We integrate the diverse drivers of conservation effects suggested by the literature into a simluation tool that is flexible enough to consider a diverse array of reasonably realistic MPA outcomes. We can use this tool to explore the range of conservation outcomes that we might expect once we allow different drivers such as movement rates and fleet dynamcis to interact together across 20,000 simulations. 

```{r}
short_term <- outcomes %>% 
  filter(years_protected <= 15, years_protected >= 0)


small_effect <- round( 100* median(short_term$mpa_effect[short_term$mpa_size <= 0.25 & short_term$depletion <= 0.5]))

min_small_effect <- round( 100* min(short_term$mpa_effect[short_term$mpa_size <= 0.25 & short_term$depletion <= 0.5]))

max_small_effect <- round( 100* max(short_term$mpa_effect[short_term$mpa_size <= 0.25 & short_term$depletion <= 0.5]))

medium_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.5 & short_term$depletion <= 0.75]))

big_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.5 & short_term$depletion > 0.75]))
```


The median simulated equilibrium regional effect, expressed as the percentage change in total biomass with the MPAs and without the MPAs, was `r round(100*median(eqo$mpa_effect))`%, with a min of `r round(100*min(eqo$mpa_effect))`% and a max of over `r round(100*max(eqo$mpa_effect))`%. Increases of over 10% took approximately 10 years to achieve for 50% of the simulated fisheries (Fig.\@ref(fig:trend-effect)). Two of the most critical factors affecting the direction and magnitude of regional consevation effects are the degree of overfishing and the size of the MPA. For cases where "small" MPAs (smaller than 25% of the range) were implemented in relatively unexploited fisheries (depletion < 50%), the median regional effect was `r small_effect`%. For moderate depletion (50% to 75%), MPA sizes from 1% to 50% produced median MPA effects of `r medium_effect`%, while for depletion above 75% the median regional effect from was `r big_effect`% (Fig.\@ref(fig:f-and-s-plot)-A). The median regional effect increased with both MPA size and depletion, but it is important to note that the ranges around these median values are extremely wide (Fig.\@ref(fig:f-and-s-plot)-B). Broadly, the simulation results show that integrating across a broad set of states of nature defined by theoretical drivers of MPA effects, under most simulations the MPAs produced positive effects, though smaller effects were much more likely than large effects. In some instances, the MPAs actually resulted in net regional conservation losses. While factors such as MPA size and pre-MPA depletion are critical drivers, even controlling for these traits a wide range of outcomes are still possible (Fig.\@ref(fig:f-and-s-plot)).

XX
The percentage change in biomass with and without MPAs is most analogous to the effects that can (in theory) be estimated by our identification strategy (the percentage difference in the density of targeted species relative to the non-targeted species pre and post MPA). However, the percentage change in biomass is a somewhat misleading metric from the perspective of meaningful conservation outcomes. Take for example an extremely depleted scenario where without MPAs a fishery is left with only 2 kg of fish. Suppose then that an MPA brings us up to 6 kg of fish, and that the unfished biomass in this fishery is 1,000kg. While the MPA has produced a 200% increase in fish biomass, this increase is relatively inconsequential given the scale of the population (we have only recovered 0.4% of the unfished biomass), and likely to be very challenging to detect in a real ecological system.
XX

To reflect this, we can repeat the analyses in Fig.\@ref(fig:trend-effect)-\@ref(fig:f-and-s-plot), but now expressing the change in biomass with and without MPAs as a percentage of unfished biomass for that fishery. Through this metric, we see median equilibrium effect sizes of `r round(100*median(eqo$pop_effect))`% (Fig. \@ref(fig:pop-trend-effect)). Over the 15 year time horizon, our simulations find that MPAs less than 25% produced a median effect of near `r round(100*median(short_term$pop_effect))`% (Fig. \@ref(fig:pop-f-and-s-plot)-A), but again with a wide range around the simulated outcomes outcomes, from `r round(100*min(short_term$pop_effect))`% to 100% (Fig. \@ref(fig:pop-f-and-s-plot)-B). 

Negative conservation outcomesfrom MPAs were primarily driven by scenarios in which the simulated fleet followed a "constant catch" strategy, especially when both depletion and MPA size were in the 25-50% range (Fig.\@ref(fig:fleet-plot)). Under the constant-catch fleet model fishing communities seek to catch the same amount regardless of the presence of an MPA. While a constant catch greater than MSY is not possible over the long-term under the confines of this simulation framework (the population would crash), over the short-term a constant-catch scenario is not implausible. Subsistence fisheries may conform to a constant-catch style policy over the short-term, as they seek to ensure the nutritional needs of their community. More industrial fisheries may have pre-arranged purchase orders for levels of catch. Quota managed fisheries may maintain relatively static quota levels until new stock assessments can be completed. The key outcome of these results is that the regional conservation effects of an MPA are critically dependent on fisheries management institutions outside the protected areas; an MPA that would provide large benefits under open-access dynamics may actually harm conservation in a constant-catch scenario. 

```{r example, fig.cap = "Example trajectories of biomass with and without MPAs under a range of MPA sizes (A), and resulting MPA conservation effect (B)", include=FALSE}

trajectories <-  processed_grid %>% 
  mutate(rounded_f = plyr::round_any(f_v_m, 0.5)) %>% 
  mutate(rounded_mpa = plyr::round_any(mpa_size, 0.25)) %>% 
  filter(rounded_f == 1.5) %>% 
  select(experiment, rounded_mpa, mpa_effect) %>% 
  unnest() %>% 
  group_by(year, rounded_mpa) %>% 
  summarise(mean_no_mpa = mean(`no-mpa`), 
            mean_wi_mpa = mean(`with-mpa`)) %>% 
  ungroup() %>% 
  mutate(mean_effect = (mean_wi_mpa - mean_no_mpa) / mean_no_mpa) %>% 
  mutate(years_protected = year - 45)

net_effect_plot <- trajectories %>% 
  filter(years_protected >= -5) %>% 
  ggplot(aes(years_protected, mean_effect, color = factor(rounded_mpa))) + 
  geom_line(size = 1.5) + 
  scale_color_viridis_d(labels = paste0(seq(0,100, by = 25),"%"),
                        name = "") + 
  scale_y_continuous(position = "right", labels = percent, name = "Conservation Effect") + 
  labs(x = "Years Protected", caption = "Facets (A) and color (B) are % of region in MPAs")

labelfoo <- function(x){
  paste0(100*as.numeric(x),"%")
  
}


traj_plot <- trajectories %>% 
  filter(rounded_mpa %in% c(0.25,075)) %>% 
  select(-mean_effect) %>% 
  gather(world, biomass, mean_no_mpa, mean_wi_mpa) %>% 
  mutate(years_protected = year - 45) %>% 
  ungroup() %>% 
  filter(years_protected >= -5) %>% 
  ggplot(aes(years_protected, biomass, color = world)) + 
  geom_line(size = 1.5, show.legend = FALSE) + 
  facet_wrap(~rounded_mpa, labeller = labeller(rounded_mpa = labelfoo)) + 
  scale_color_manual(values = c("#009ACD", "#548B54") ,name = element_blank(),
                     labels = c("Without MPA","With MPA")) + 
  labs(y = "Biomass", x = "")


example_mpa_plot <- traj_plot + 
  net_effect_plot  +  
  plot_annotation(tag_levels = "A") +
  plot_layout(ncol = 2,
              nrow = 1,widths = c(1.5, 1)) & 
  theme(
    plot.margin = unit(c(.2, .2, .2, .2), unit = "lines"),
    panel.spacing = unit(.2, unit = "lines"),
    legend.position = "top",
    strip.text = element_text(size = 2)
    )

# ggsave("figs/example.pdf",traj_plot, width = 4, height = 4, units = "in") 
  
pdf(file =  "figs/example.pdf",width = 4, height = 3, useDingbats = TRUE)
print(example_mpa_plot)
dev.off()

  
```


Figure \@ref(fig:example) is an example of putting a frog in somewhere. 


\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{figs/example.pdf}
\caption{Placeholder image of a frog with a long example caption to show justification setting.}
\label{fig:example}
\end{figure}


Fig.\ref{fig:example} shows a thing

figure 2.4 showing distribution of effects after XX years (15?) 

```{r conservation-effect, fig.cap = "Median (A) and range (B) regional MPA conservation effect (expressed as percent of unfished biomass) after 15 years of protection across a range of pre-MPA depletions and MPA sizes", include = FALSE}


pop_depletion_plot <- outcomes %>%
  filter(years_protected == 15) %>%
  ggplot() +
  geom_bin2d(aes(depletion, pop_effect), binwidth = c(.05, .05),
             show.legend = FALSE) +
  scale_fill_viridis(
  option = "A",
  trans = "log10",
  guide = guide_colorbar(frame.colour = "black",
  frame.linewidth = 1,),
  name = "Median Effect"
  ) +
  scale_x_continuous(labels = percent, name = "Pre-MPA Depletion") + 
  scale_y_continuous(labels = percent, name = "")


pop_size_plot <- outcomes %>%
  filter(years_protected == 15) %>%
  ggplot() +
  geom_bin2d(aes(mpa_size, pop_effect), binwidth = c(.05, .05), show.legend = TRUE) +
  scale_fill_viridis(
  option = "A",
  trans = "log10",
  guide = guide_colorbar(frame.colour = "black",
  frame.linewidth = 1,),
  name = "# of Sims"
  ) +
  scale_x_continuous(labels = percent, name = "Range in MPA") + 
    scale_y_continuous(labels = percent, name = "Conservation Effect") + 
  theme(legend.position = "right")


 pop_facet_effect_plot <- outcomes %>%
  filter(years_protected == 15) %>%
   select(pop_effect, mpa_size, depletion) %>% 
   gather(measure, value, -pop_effect) %>% 
   ggplot() +
   geom_bin2d(aes(value, pop_effect), binwidth = c(.05, .05), show.legend = TRUE) +
  scale_fill_viridis(
  option = "A",
  trans = "log10",
  guide = guide_colorbar(frame.colour = "black",
  frame.linewidth = 1,barheight = unique(10, "lines")),
  name = "# of Sims"
  ) + 
   facet_wrap(~measure, labeller = labeller(measure = facet_labels), strip.position = "bottom") + 
   scale_x_percent(name = "") + 
   scale_y_percent(name = "Conservation Effect")


  pop_depletion_and_size_plot <- outcomes %>% 
    filter(years_protected >= 0) %>%
    group_by(experiment) %>% 
    mutate(depletion = plyr::round_any(depletion[years_protected == 0], .05),
           mpa_size = plyr::round_any(mpa_size, .05)) %>% 
    filter(years_protected == 15) %>% 
    group_by(depletion, mpa_size) %>% 
    summarise(median_mpa_effect = median(pop_effect)) %>% 
    ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) + 
    geom_tile() + 
    # geom_contour(aes(z = median_mpa_effect)) +
    scale_fill_viridis(labels = percent,
                       guide = guide_colorbar(frame.colour = "black",
                                              frame.linewidth = 1,
                                              barheight = unit(10,"lines")), 
                       name = "Median Effect") +
  labs(x = "Pre-MPA Depletion", y = "Range in MPA") + 
    scale_y_continuous(labels = percent) + 
    scale_x_continuous(labels = percent)
  
  pop_combo_plot <-
    (pop_depletion_and_size_plot + labs(title = "A")) + ((pop_size_plot + labs(title = "B")) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) & theme(
    plot.margin = unit(c(0, 0, 0, 0), units = "lines"),
    axis.text.x = element_text(size = 8),
    legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
    axis.text.y = element_text(size = 10)
    )
    
  
  pdf(file =  "figs/conservation-effect.pdf",width = 8, height = 5, useDingbats = TRUE)
print(pop_combo_plot)
dev.off()
  
```


Hello there, what's up

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{figs/conservation-effect.pdf}
\caption{Placeholder image of a frog with a long example caption to show justification setting.}
\label{fig:conservation-effect}
\end{figure*}


Fishery Effects {#fisheries .unnumbered}
-------------------

Mention problems with some papers on this topic

We express the fishery effects as a percentage of maximum sustainable yield (MSY) gained or lost in the world with MPAs relative to the world without MPAs. The fishery effect now reflects the percentage of MSY gained or lost as a result of MPA implementation. The overall trends are similar to the relative change in catch results, but the 15 year effects are more muted, with peak positive effects near 25% and negative effects near -75% (Fig.\@ref(fig:catch-msy-effects)). 

XX what else goes in here?

```{r fishery-effects,fig.cap = "Median (A) and range (B) MPA fishery effects, expressed as the difference in catch with and without MPAs  as a proportion of MSY, after 15 years of protection across a range of pre-MPA depletions and MPA sizes", include = FALSE}

msy_depletion_plot <- fishery_outcomes %>%
  filter(years_protected == 15) %>%
  ggplot() +
  geom_bin2d(aes(depletion, msy_effect), binwidth = c(.05, .05),
             show.legend = FALSE) +
  scale_fill_viridis(
  option = "A",
  trans = "log10",
  guide = guide_colorbar(frame.colour = "black",
  frame.linewidth = 1,),
  name = "Median Effect"
  ) +
  scale_x_continuous(labels = percent, name = "Pre-MPA Depletion") + 
  scale_y_continuous(labels = percent, name = "Fishery Effect")


msy_size_plot <- fishery_outcomes %>%
  filter(years_protected == 15) %>%
  ggplot() +
  geom_bin2d(aes(mpa_size, msy_effect), binwidth = c(.05, .05), show.legend = TRUE) +
  scale_fill_viridis(
  option = "A",
  trans = "log10",
  guide = guide_colorbar(frame.colour = "black",
  frame.linewidth = 1,),
  name = "# of Sims"
  ) +
  scale_x_continuous(labels = percent, name = "Range in MPA") + 
    scale_y_continuous(labels = percent, name = "Fishery Effect") + 
  theme(legend.position = "right")

msy_facet_effect_plot <- fishery_outcomes %>%
  filter(years_protected == 15) %>%
   select(msy_effect, mpa_size, depletion) %>% 
   gather(measure, value, -msy_effect) %>% 
   ggplot() +
   geom_bin2d(aes(value, msy_effect), binwidth = c(.05, .05), show.legend = TRUE) +
  scale_fill_viridis(
  option = "A",
  trans = "log10",
  guide = guide_colorbar(frame.colour = "black",
  frame.linewidth = 1,barheight = unique(10, "lines")),
  name = "# of Sims"
  ) + 
   facet_wrap(~measure, labeller = labeller(measure = facet_labels), strip.position = "bottom") + 
   scale_x_percent(name = "") + 
   scale_y_percent(name = "Fishery Effect")


  msy_depletion_and_size_plot <- fishery_outcomes %>% 
    filter(years_protected >= 0,!is.na(msy)) %>%
    group_by(experiment) %>% 
    mutate(depletion = plyr::round_any(depletion[years_protected == 0], .05),
           mpa_size = plyr::round_any(mpa_size, .05)) %>% 
    filter(years_protected == 15) %>% 
    group_by(depletion, mpa_size) %>% 
    summarise(median_mpa_effect = median(msy_effect, na.rm = T)) %>% 
    ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) + 
    geom_tile() + 
    # geom_contour(aes(z = median_mpa_effect)) +
    scale_fill_gradient2(midpoint = 0,
                         low = "tomato",
                         high = "steelblue",
                         mid = "white",
                         labels = percent,
                       guide = guide_colorbar(frame.colour = "black",
                                              frame.linewidth = 1,
                                              barheight = unit(10,"lines")), 
                       name = "Median Effect") +
  labs(x = "Pre-MPA Depletion", y = "Range in MPA") + 
    scale_y_continuous(labels = percent) + 
    scale_x_continuous(labels = percent)
  
  msy_fishery_combo_plot <-
    (msy_depletion_and_size_plot + labs(title = "A")) + ((msy_size_plot + labs(title = "B")) / msy_depletion_plot)  + plot_layout(widths = c(1.5, 1)) & theme(
    plot.margin = unit(c(0, 0, 0, 0), units = "lines"),
    axis.text.x = element_text(size = 8),
    legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
    axis.text.y = element_text(size = 10)
    )
    
pdf(file =  "figs/fishery-effect.pdf",width = 8, height = 5, useDingbats = TRUE)
print(msy_fishery_combo_plot)
dev.off()
  
```


Estimating Conservation Effects {#estimating .unnumbered}
-------------------

Theory and simulation testing indicate then that while the regional effects of MPAs can range from strongly negative to highly positive, the bulk of scenarios produced regional conservation effects within the range of  0-10% after fifteen years of protection. Given this, our expectation is that the "true" regional effect will likely be challenging to isolate from the variation of natural systems in and the observation error inherent to any MPA monitoring program. We use data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring of the Channel Islands National Marine Sanctuary to test our ability to detect the regional effect of MPAs in a real world context. PISCO conducts visual underwater SCUBA surveys at a variety of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands. At the rawest level, the data are counts of finfish in 2cm length bins along a 30m x 2m transect at various sites and depths. These length bins are converted to biomass, and then biomass densities, by converting length to weights using available allometric data and dividing by the transect area. Our goal is to estimate the effect of the MPAs on these densities of fish throughout the Channel Islands. 

Our identification strategy is to use non-targeted species as our control for unaccounted for environmental trends before and after MPA implementation (which occurred in 2003). The model estimates the difference in the trends between targeted and non-targeted species pre- and post-MPA. We hypothesize that there should be no difference in pre-MPA trends. We fit this model using a hierarchical mixed-effect framework using Template Model Builder [TMB, @Kristensen2016] in R [@RCoreTeam2018]. The model consists of three levels, the first (starting from the "bottom") being transect-level densities of fish species observed by PISCO, which are standardized into a unit-less index of biomass abundance (which we will refer to as an abundance index from now now), accounting for both probability of detection and expected density as a result of changes in both abundance and covariates such as observer skill [see @Maunder2004]. For the second stage, we break the abundance indices into targeted and non-targeted species (per the classifications in the PISCO data), and estimate the mean trend of each group (targeted and non-targeted) over time. In the third step, we estimate the difference in the mean trend between the targeted and non-targeted fishes, which under the right set of circumstances should reflect the causal effect of the MPAs on the outcome of interest (in this case regional biomass density of targeted fishes). Xx DID citation All three of these steps are integrated into the same estimation model, in order to propagate uncertainty through the model correctly.

The proposed identification strategy serves to control for some unobserved factors influencing densities of targeted and non-targeted species, but is unlikely to account for all of them. Before examining regression results, we can graphically examine the trends in mean densities for targeted and non-targeted species over time. We centered and scaled the mean annual densities for each species included in the analysis in order to compare the trends across species groups. Grouping the species by targeted and non-targeted status, we see evidence of pre-treatment parallel trends in the abundance indices, and of a divergence post MPA implementation. Beginning in around 2007 abundances of targeted species appear to start increasing faster than non-targeted species. However, from 2012 onward abundances of targeted species appear to be declining relative to the trend in the non-targeted species. Not controlling for any other factors that may affect fish abundances, the data suggests a possible increase in targeted species abundance (relative to the "control" trend of the non-targeted species) at first, followed by a decrease in the most recent years (Fig.\@ref(fig:raw-trend)). 




And here's a map of the channel islands

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
\caption{Placeholder image of a frog with a long example caption to show justification setting.}
\label{fig:channel-islands}
\end{figure}



```{r ci-map, fig.cap = "Map of study region and sampling locations. Shaded polygons indicate location of MPAs. Points represent sampling locations, and color indicates the number of observations recorded at a given point", include = FALSE}

ci_map <- pisco_data %>% 
  left_join(site_data, by = c("site","side")) %>% 
  filter(is.na(eventual_mpa) == F) %>% 
  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
         classcode %in% top_species) %>% 
  group_by(lat_wgs84, lon_wgs84, site, region, eventual_mpa) %>% 
  count()



ci_map <-  ci_map %>%
  dplyr::mutate(geometry = purrr::map2(lon_wgs84, lat_wgs84, ~ sf::st_point(x = c(.x, .y), dim = 'XY'))) %>%
  ungroup() %>%
  mutate(geometry = sf::st_sfc(geometry, crs = 4326)) %>%
  sf::st_sf()


bbox <- sf::st_bbox(ci_map)

pisco_ci_map_plot <-  ggmap(channel_islands) +
  geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
  geom_sf(data = ci_map,inherit.aes = FALSE,
  aes(color = n),
  alpha = 0.75,
  size = 1) +
  coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
  ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .4)) +
  scale_color_gradient(low = "white", high = "orangered",
  guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
  barheight = unit(13, "lines")),
  name = "Samples") + 
  ggthemes::theme_map()

pdf(file =  "figs/channel-islands.pdf",width = 8, height = 5, useDingbats = TRUE)
print(pisco_ci_map_plot)
dev.off()


```

Or more things

Results {#results .unnumbered}
------

```{r did-plot, fig.cap = "Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (i.e. integrated across inside and outside of MPAs). MPAs are implemented in 2003 (red dashed line). Estimates are from a regression on log(abundance index), so estimated effects roughly correspond to percentage changes", include = FALSE}

did_plot <- did_plot + 
  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = percent)

pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
print(did_plot)
dev.off()

```



Discussion {#discussion .unnumbered}
------


Despite the vast amount of rigorous data collection before and after, our identification strategy was unable to identify a clear signal of the effect of MPAs on the regional densities of targeted species. Our simulation analysis indicates though that we should not be surprised by this. The Channel Islands MPAs cover approximately 21% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis what evidence we have does not suggest they as a group they are heavily overfished. Our simulation analysis would suggest then that the percentage difference in densities of targeted species with and without MPAs should be on the smaller end (10% or less), and therefore be challenging to detect given the natural variation of marine ecosystems and the error inherent in visual survey programs such as those the PISCO data used here. What then does this suggest for the design and monitoring of MPAs?

Real world-policy making is inherently an exercise in utilizing best available theory, experience, and modeling to make decisions that are often difficult or practically impossible to empirically test. Despite our best efforts we are unlikely to ever truly "know" the effect of our efforts to mitigate climate change, but must instead rely on comparisons to best available modeling outcomes to understand how effective our policies have been. Similarly, given the complexities of marine systems much of our decisions on MPA design will have to be based on effective modeling. While we are hardly the first to point out that bio-economic models are a critical tool for MPA design, our results help indicate a minimum floor of model complexity to provide candid assessments of regional MPA effects. While factors such as MPA size and degree of depletion are especially strong drivers, for all but the most extreme cases of each a wide array of effects, from negative to highly positive,  are possible based on the complex interaction of factors such as fleet dynamics, movement rates, and recruitment timing. Confronting these interactions by considering the likely parameter space for a given region is a critical step then in understanding what likely regional effects of MPAs are. While models such as ours do require large numbers of parameters that may be challenging to obtain, our results show that working with communities to confront these uncertainties is preferable to sweeping them under the rug in favor of simpler models that are easier to parameterize but miss details that our results show can have dramatic effect on expected outcomes. Modeling effort such as this can then provide stakeholders with some idea of the range of regional effect sizes that might be expected for a given MPA network design, and in doing so design monitoring programs targeted at the species and fleets that modeling suggests may provide the clearest indication of MPA mediated effects. For cases where bio-economic modeling suggests small potential for MPA driven regional density effects, monitoring efforts can be targeted around detecting potential negative effects should they arise, i.e. evidence that the model is wrong, rather than exerting massive amounts of effort on what theory and modeling suggest may be a small effect size. 

We focus mostly on regional conservation gains in this paper. However, fisheries spillover is often another important factor to consider (i.e. are fisheries better off with the MPAs than they would have been without them). The fishery benefits of MPAs are just as (and likely more) intensely debated than the regional conservation outcomes [@Hilborn2004; @Roberts2001;@Sala2018b;@Hilborn2018]. We only address fisheries affects briefly in this study, but our results highlight important tradeoffs and synergies between conservation and fishery spillover effects of MPAs. The good news from a fisheries perspective is also fairly obvious: Both the regional conservation and fishery benefits are expected to be greatest when the fished population is in an extremely depleted state pre-MPA, even over a 15 year time horizon, even for larger MPAs (though further work is needed to compare MPAs to alternative fisheries management strategies in these cases). For cases where a valuable and formerly abundant species is overfished, a large MPA may then provide large conservation and fishery gains for that species, while potentially having smaller impacts on other less depleted species. Our simulation results also do identify though a large parameter space where MPAs create tradeoffs between moderate conservation gains and moderate fishing losses. This type of projection analysis can help managers consider where in this space they may be. The most critical point with regards to conservation and fishery effects from our simulation analysis is that the conservation or fishery effects of MPAs cannot be reliably estimated without some knowledge and consideration of the dynamics of the fishing fleet outside the MPAs: over the short-term open access vs constant catch dynamics can make the difference between a substantial conservation and fisheries win to a more depleted stock with more expensive fishing. 

MPAs are an important part of the marine resource management toolbox. Under ideal circumstances they can protect both individual species and ecosystem linkages, safeguard crucial habitat, and support local economies through tourism and fishing opportunities. However, our results show that the regional conservation and fishery benefits we should expect from MPAs are highly variable, and while we cannot assign probabilities to our simulated states of nature, we find that there are more opportunities for smaller effect sizes than large, especially in the short-term. Most importantly, our results highlight the critical importance of explicitly modeling the links between the biological effects of MPAs and the ways that humans respond to them. Large-scale empirical evidence supports our results that accurately predicting the effects of MPAs depends on understanding the human context in which they find themselves [@Cinner2018]. While this is far from the first effort at simulating MPAs, our model fills an important niche between more focused theoretical models that address a few drivers of MPA performance at a time but do not address complicated linkages between drivers, and more site-specific models that provide best available local results but are challenging to scale to different systems. This intermediate complexity model allows us to simulate tens of thousands of fisheries containing most of the key drivers of MPA performance identified by theory. 

This process provides a unique survey of the likely regional effects of MPAs to fisheries and conservation, and places our empirical assessment of the regional effects of the Channel Islands MPAs in context. The Channel Islands are an intensely studied system, and the challenge of identifying a clear regional effect of the network of MPAs placed there in 2003 may seem surprising. However, our results show that in fact a smaller effect size, from the perspective of regional conservation gains, is to be expected in this system, and therefore the true effect will be very challenging to separate from environmental noise. The solution then though is not to give up on detecting effects, but rather to shift focus from identifying a specific effect size and instead use simulation analysis to appropriately set expectations for conservation and fishing stakeholders, and design monitoring programs around the species and situations that serve as effective indicators of the ability of an MPA network to achieve its objectives. 


Methods {#methods .unnumbered}
------

Supporting Information (SI) {#supporting-information-si .unnumbered}
---------------------------


The main text of the paper must stand on its own without the SI. Refer
to SI in the manuscript at an appropriate point in the text. Number
supporting figures and tables starting with S1, S2, etc. Authors are
limited to no more than 10 SI files, not including movie files. Authors
who place detailed materials and methods in SI must provide sufficient
detail in the main text methods to enable a reader to follow the logic
of the procedures and results and also must reference the online
methods. If a paper is fundamentally a study of a new method or
technique, then the methods must be described completely in the main
text. Because PNAS edits SI and composes it into a single PDF, authors
must provide the following file formats only.

### SI Text {#si-text .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference).

### SI Figures {#si-figures .unnumbered}

Provide a brief legend for each supporting figure after the supporting
text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or
GIF format; figures may not be embedded in manuscript text. When saving
TIFF files, use only LZW compression; do not use JPEG compression. Do
not save figure numbers, legends, or author names as part of the image.
Composite figures must be pre-assembled.

### 3D Figures {#d-figures .unnumbered}

Supply a composable U3D or PRC file so that it may be edited and
composed. Authors may submit a PDF file but please note it will be
published in raw format and will not be edited or composed.

### SI Tables {#si-tables .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference); include only one
table per file. Do not use tabs or spaces to separate columns in Word
tables.

### SI Datasets {#si-datasets .unnumbered}

Supply Excel (.xls), RTF, or PDF files. This file type will be published
in raw format and will not be edited or composed.

### SI Movies {#si-movies .unnumbered}

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{figs/did_plot.pdf}
\caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (i.e. integrated across inside and outside of MPAs). MPAs are implemented in 2003 (red dashed line). Estimates are from a regression on log(abundance index), so estimated effects roughly correspond to percentage changes.}
\label{fig:did}
\end{figure}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
(wmv), animated GIF (gif), or MPEG files and submit a brief legend for
each movie in a Word or RTF file. All movies should be submitted at the
desired reproduction size and length. Movies should be no more than 10
MB in size.

### Still images {#still-images .unnumbered}

Authors must provide a still image from each video file. Supply TIFF,
EPS, high-resolution PDF, JPEG, or GIF files.

### Appendices {#appendices .unnumbered}

PNAS prefers that authors submit individual source files to ensure
readability. If this is not possible, supply a single PDF file that
contains all of the SI associated with the paper. This file type will be
published in raw format and will not be edited or composed.


\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper. 

\subsection*{Subsection for Method}
Example text for subsection.
}

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak


