---
title: "Developing MPA sims"
output: html_notebook
---

```{r}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# Summary: Load things you need
rm(list = ls())
set.seed(123)
library(tidyverse)
library(forcats)
library(modelr)
library(stringr)
library(rstan)
library(rstanarm)
library(car)
library(AER)
library(broom)
library(viridis)
library(scales)
library(spasm)
library(rfishbase)

demons::load_functions()

```


Goal here is to explore simulation strategies for demonstrating the conservation effects of MPAs, to decide what set you are going to include in the final chapter/paper. 

The basis for this is the `spasm` package. As a starting point, let's explore what just messing with biology and starting conditions does. 

As a second point, focus on changing fleet dynamics and response

As a final point, mess with species composition and depletion assemblage


# Biological Effects

To start, let's hunt over ranges of age at maturity, steepness, adult movement, larval movement, and density dependence forms, and natural mortality, and fishing pressure

```{r, eval = F}

sim_grid <- expand.grid(max_age = seq(5,20, length.out = 2),
                        age_mat = seq(1,.75, length.out = 3),
                        steepness = seq(0.3,.9, length.out = 2),
                        adult_movement = seq(1,100, length.out = 3),
                        larval_movement = seq(1,100, length.out = 3),
                        density_dependence_form = 1:5,
                        vbk = seq(1e-2, .5, length.out = 2))

sim_grid <- sim_grid %>% 
  mutate(fish = pmap(list(max_age = max_age,
                          age_mat = max_age * age_mat,
                          steepness = steepness,
                          adult_movement = adult_movement,
                          larval_movement = larval_movement,
                          vbk = vbk,
                          density_dependence_form = density_dependence_form), create_fish,
                     scientific_name = 'fakeish fishis',
                     query_fishbase = F,
                     linf = 100,
                     t0 = -.2,
                     weight_a = 1e-4,
                     weight_b = 3))

sim_grid <- sim_grid %>% 
  mutate(eq_f = rlnorm(dim(.)[1], log(0.5)) * map_dbl(fish,'m'),
    fleet = pmap(list(fish = fish, eq_f = eq_f), 
                      create_fleet,
                              length_50_sel = 1,
                              length_95_sel = 2))
a <- proc.time()

sim_grid <- sim_grid %>% 
  mutate(simulation = pmap(list(fish = fish, fleet = fleet),
                           sim_fishery,
                           manager = create_manager(),
                           sim_years = 50,
                           burn_years = 25,
                            num_patches = 100))

proc.time() - a
  
plot_simulation <- function(sim){

sim %>%
    group_by(year) %>%
    summarise(ssb = sum(ssb), prop_mpa = mean(mpa)) %>%
    ggplot(aes(year,ssb, fill = prop_mpa)) +
    geom_point(shape = 21) +
    scale_fill_viridis(labels = percent, name = '% MPA') +
    xlab('Year') +
    ylab('SSB')

}

sim_grid <- sim_grid %>%
  mutate(sim_plot = map_plot(simulation,plot_simulation)) %>% 
  select(-fish,-fleet)

trelliscope(sim_grid, name = 'testing')

```

Well that was a giant waste of time. Really hard to interpret anything out of that. Well, no, it does tell you broadly that there's not a lot interesting going on with the biological parameters per say, though their interaction with other things still might be important. 

Big problem is the lack of a counterfactual. You need a function that is similar to your current one, but shows trajectories with and without MPA, which is really what you care about

New approach. Let's pick a few case scenarios and look at them more in depth. 


```{r, eval = F}

comp_foo <- function(fish, fleet, year_mpa,mpa_size,sim_years, num_patches,
                     burn_years){
  
  
  no_mpa <- sim_fishery(fish = fish, fleet = fleet, manager = create_manager(year_mpa = year_mpa, mpa_size = 0), sim_years = sim_years, num_patches = num_patches,
                        burn_years = burn_years) %>% 
    mutate(experiment = 'no-mpa')
  
    wi_mpa <- sim_fishery(fish = fish, fleet = fleet, manager = create_manager(year_mpa = year_mpa, mpa_size = mpa_size), sim_years = sim_years, num_patches = num_patches,
                        burn_years = burn_years) %>% 
      mutate(experiment = 'with-mpa')
    
    outcomes <- no_mpa %>% 
      bind_rows(wi_mpa) %>% 
      group_by(year,experiment) %>% 
      summarise(ssb = sum(ssb), percent_mpa = mean(mpa))
    

  
} 

year_mpa <- sample(1:25, 200, replace = T)

mpa_size <- runif(200,0,1)

sim_grid_outcomes <- sim_grid %>% 
  sample_n(200) %>% 
  mutate(year_mpa = year_mpa,
         mpa_size = mpa_size) %>% 
  mutate(outcomes = pmap(list(fish = fish, fleet = fleet,
                              year_mpa = year_mpa,
                              mpa_size = mpa_size),
                           comp_foo,
                           sim_years = 50,
                           burn_years = 25,
                            num_patches = 100
                         ))
  


plot_foo <- function(outcomes, year_mpa){
  
    min_year <- min(outcomes$year)
    
    mpa_size <-  max(outcomes$percent_mpa)
   
     outcomes %>% 
      filter(year > (min_year + year_mpa - 3) & year < (min_year + year_mpa + 10)) %>% 
      ggplot(aes(year,ssb)) + 
       geom_line(aes(color = experiment)) +
      geom_point(size = 2,shape = 21,aes(fill = experiment)) +
      geom_vline(aes(xintercept = year_mpa + min_year - 1)) +
      theme_ipsum() +
       labs(caption = paste0(100*mpa_size,'% MPA'))
}

sim_grid_outcomes <- sim_grid_outcomes %>%
  mutate(sim_plot = map2_plot(outcomes,year_mpa,plot_foo)) %>% 
  mutate(eq_f = map_dbl(fleet,'eq_f')) %>% 
  select(-fish,-fleet)

trelliscope(sim_grid_outcomes)

sim_grid_outcomes$sim_plot[[1]]
plot_foo(sim_grid_outcomes$outcomes[[1]])
```


OK, incredibly clear that you can't get an MPA mediated conservation decrease purely as a function of life history, fishing mortality, and stage of MPA implementation. Basically no what, wh The only thing you haven't really explored yet is a "rebuilding" fishery, but I'm pretty sure that isn't the culprit. 

Seems like the most logical thing to start building in here is fleet dynamics. Basically, in order for your hypothesis to work, at least at first the relative increase in F has to be greater than the spillover benefit accruing at first. Might actually be mathematically impossible. 

So, let's add in two factors first: open access dynamics, and fishing the line. 

Hmmm neither of those really do anything. At the end of the day, the only way that MPAs can decrease biomass is if it causes a net increase in F in at least the early years. How can F increase when you're shutting off fishing? If you think of F as C/B, then the proportion of catch to biomass has to go up. If you tried a constant catch policy that doesn't seem like it would do it, since B should go up a bit right away post MPA, so, F should decrease marginally on net at least at first. 

You should code it up quickly to see if it does anything, and after that put a pin in things. What do you do if you can't come up with a defensible theory for why biomass would go down? You've covered constant effort, open access, constant catch.... what else is there? Blue paradox? A sort of fleet "over-reaction" to the MPA in response to perceived loss later? 

# Test fleet effects

```{r}

library(tidyverse)
library(spasm)
library(rfishbase)

comp_foo <- function(fish, fleet, year_mpa,mpa_size,sim_years, num_patches,
                     burn_years){
  
  
  no_mpa <- sim_fishery(fish = fish, fleet = fleet, manager = create_manager(year_mpa = year_mpa, mpa_size = 0), sim_years = sim_years, num_patches = num_patches,
                        burn_years = burn_years) %>% 
    mutate(experiment = 'no-mpa')
  
    wi_mpa <- sim_fishery(fish = fish, fleet = fleet, manager = create_manager(year_mpa = year_mpa, mpa_size = mpa_size), sim_years = sim_years, num_patches = num_patches,
                        burn_years = burn_years) %>% 
      mutate(experiment = 'with-mpa')
    
    outcomes <- no_mpa %>% 
      bind_rows(wi_mpa) %>% 
      group_by(year,experiment) %>% 
      summarise(ssb = sum(ssb), percent_mpa = mean(mpa),
                catch = sum(biomass_caught),
                profits = sum(profits),
                effort = sum(effort))
    

  
} 

sim_grid <- expand.grid(max_age = 20,
                        age_mat = 0.5,
                        steepness = seq(0.3,.9, length.out = 2),
                        adult_movement = seq(1,100, length.out = 3),
                        larval_movement = seq(1,100, length.out = 3),
                        density_dependence_form = 1:5,
                        vbk =.2,
                        target_catch = seq(1,1000, length.out = 10),
                        mpa_size = seq(0,1, length.out = 4))

sim_grid <- sim_grid %>%
  mutate(fish = pmap(list(max_age = max_age,
                          age_mat = max_age * age_mat,
                          steepness = steepness,
                          adult_movement = adult_movement,
                          larval_movement = larval_movement,
                          vbk = vbk,
                          density_dependence_form = density_dependence_form), create_fish,
                     scientific_name = 'fakeish fishis',
                     query_fishbase = F,
                     linf = 100,
                     t0 = -.2,
                     weight_a = 1e-4,
                     weight_b = 3))

sim_grid <- sim_grid %>%
  mutate(
    fleet = pmap(list(target_catch = target_catch, fish = fish), create_fleet,
                 fleet_model = 'constant-catch', initial_effort = 0))



year_mpa <- 25

sim_grid_outcomes <- sim_grid %>%
  mutate(year_mpa = year_mpa)  %>%
  sample_n(1) %>%
  mutate(outcomes = pmap(list(fish = fish, fleet = fleet,
                              year_mpa = year_mpa,
                              mpa_size = mpa_size),
                         comp_foo,
                         sim_years = 50,
                         burn_years = 25,
                         num_patches = 100
  ))



plot_foo <- function(outcomes, year_mpa){

  min_year <- min(outcomes$year)

  mpa_size <-  max(outcomes$percent_mpa)

  outcomes %>%
    filter(year > (min_year + year_mpa - 3) & year < (min_year + year_mpa + 10)) %>%
    ggplot(aes(year,ssb)) +
    geom_line(aes(color = experiment)) +
    geom_point(size = 2,shape = 21,aes(fill = experiment)) +
    geom_vline(aes(xintercept = year_mpa + min_year - 1)) +
    theme_ipsum() +
    labs(caption = paste0(100*mpa_size,'% MPA'))
}

sim_grid_plots <- sim_grid_outcomes %>%
  mutate(sim_plot = map2_plot(outcomes,year_mpa,plot_foo)) %>%
  mutate(eq_f = map_dbl(fleet,'eq_f')) %>%
  select(-fish,-fleet,-outcomes)


a %>% 
  gather('metric','value',ssb,catch,profits, effort) %>% 
  ggplot(aes(year,value)) + 
  geom_vline(aes(xintercept = year_mpa + min(year) - 1), linetype = 2) +
  geom_line(aes(color = experiment)) + 
  facet_wrap(~metric, scales = "free_y")

trelliscope(sim_grid_plots,name = 'arg')

```


