---
title: "MLPA = More Fat Fish?"
subtitle: "Gaines Lab Meeting"
author: "Dan Ovando"
date: "2017/10/12"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message = F, warning = F, echo = F, fig.align = 'center', dev = 'svg', fig.height = 5, fig.width = 8)
```

```{r load-life}

library(sf)
library(scales)
library(rstanarm)
library(scales)
library(viridis)
library(ggmap)
library(tidyverse)
library(forcats)
library(stringr)
library(lubridate)
library(purrr)
library(hrbrthemes)
library(extrafont)
library(wesanderson)


demons::load_functions('../functions')

pres_theme <- theme_ipsum(base_size = 15)

theme_set(pres_theme)

run_name <- 'Working'

run_dir <- file.path('../results', run_name)

run_description <-
  'Model selection process, testing STAN selection'

if (dir.exists(run_dir) == F) {
  dir.create(run_dir)
}

load(file.path(run_dir,'mpa-sims.Rdata'))

run_dir <- file.path('../results', run_name)

run_description <-
  'Model selection process, testing STAN selection'

if (dir.exists(run_dir) == F) {
  dir.create(run_dir)
}

load(file = paste0(run_dir,'/did-models.Rdata'))

load(file = paste0(run_dir,'/abundance-model-data.Rdata'))

load(file = paste0(run_dir,'/model_comparisons.Rdata'))

length_data <- read_csv('../data/UCSB_FISH raw thru 2013.csv') %>%
  magrittr::set_colnames(., tolower(colnames(.)))

life_history_data <-
  read_csv('../data/VRG Fish Life History in MPA_04_08_11_12 11-Mar-2014.csv') %>%
  rename(classcode = pisco_classcode) %>%
  mutate(classcode = tolower(classcode)) %>%
  # rename(description_2 = Description) %>%
  magrittr::set_colnames(., tolower(colnames(.)))

site_data <- read_csv('../data/Final_Site_Table_UCSB.csv') %>%
  magrittr::set_colnames(., tolower(colnames(.))) %>%
  select(site,side,mpagroup, mpa_status, reserve, region, year_mpa,mpaareanm2) %>%
  unique()

length_data <- length_data %>%
  left_join(life_history_data %>% mutate(classcode = toupper(classcode)), by = 'classcode') #%>%
  # left_join(site_data, by = c('site', 'side'))

conditions_data <- length_data %>%
  group_by(site,side, year) %>%
  summarise(
    mean_temp = mean(temp, na.rm = T),
    mean_kelp = mean(pctcnpy, na.rm = T),
    mean_vis = mean(vis, na.rm = T)
  )


density_data <- read_csv('../data/ci_reserve_data_final3 txt.csv') %>%
  magrittr::set_colnames(., tolower(colnames(.))) %>%
  gather('concat.name', 'value', grep('_', colnames(.)), convert = T) %>%
  mutate(
    data.type = gsub('\\_.*', '', concat.name),
    classcode = gsub('.*\\_', '', concat.name)
  ) %>%
  mutate(value = as.numeric(value)) %>%
  spread(data.type, value) %>%
  rename(site_side = site.side)



site_coords <- density_data %>% 
  group_by(site, side) %>% 
  summarise(latitude = mean(lon.wgs84, na.rm = T),
            longitude = mean(lat.wgs84, na.rm = T))

# ggmap::qmplot(longitude,latitude, color = side, data = site_coords)


if (file.exists('../data/enso.csv')) {
  enso <- read_csv('../data/enso.csv') %>%
    group_by(year) %>%
    summarise(mean_enso = mean(enso, na.rm = T)) %>%
    mutate(lag1_enso = dplyr::lag(mean_enso,1),
           lag2_enso = dplyr::lag(mean_enso,2),
           lag3_enso = dplyr::lag(mean_enso,3),
           lag4_enso = dplyr::lag(mean_enso,4))

} else {
  scrape_enso(outdir = '../data/')

}

if (file.exists('../data/pdo.csv')) {
  pdo <- read_csv('../data/pdo.csv') %>%
    group_by(year) %>%
    summarise(mean_pdo = mean(pdo, na.rm = T)) %>%
    mutate(lag1_pdo = dplyr::lag(mean_pdo,1),
           lag2_pdo = dplyr::lag(mean_pdo,2),
           lag3_pdo = dplyr::lag(mean_pdo,3),
           lag4_pdo = dplyr::lag(mean_pdo,4))

} else {
  scrape_pdo(outdir = '../data/')

  pdo <- read_csv('../data/pdo.csv') %>%
    group_by(year) %>%
    summarise(mean_pdo = mean(pdo, na.rm = T)) %>%
    mutate(lag1_pdo = dplyr::lag(mean_pdo,1),
           lag2_pdo = dplyr::lag(mean_pdo,2),
           lag3_pdo = dplyr::lag(mean_pdo,3),
           lag4_pdo = dplyr::lag(mean_pdo,4))

}




```


# Goals of the MLPA

1. Protect the natural diversity and abundance of marine life, and the structure, function and integrity of marine ecosystems.

2. Help sustain, conserve and protect marine life populations, including those of economic value, and rebuild those that are depleted.


--

** 14 years in, did MLPA increase density of fishes in the Channel Islands? **

---

# Hypotheses

- "Classic MPA": Slow and steady increase in densities

- "Scorched Earth": Decreased densities outside, increased inside, net effect unclear

- "Ms. Frizzle": Environment drives the bus; who knows what happens


---


# Identification Strategy

- Changes in density may be due to environmental confounders (e.g. El Niño)

- Proposed solution: *Unfished species (garibaldi, señorita, etc.) are environmental "control" for fished species (sheephead, rockfish)*

- Rationale
  
  - Environment affects all species
  
  - Fishing only affects some (sort of)
  
  - Trend of unfished species post-MPA then serves as counterfactual for fished

- Key Assumptions
  
  - Parallel trends
  
  - No interactions (or net zero interactions)
  

---

class: inverse, center, middle
# Data Data Data

---

# What do we have to work with?


- Underwater visual survey transects

  - Stratfied random sampling inside and outside eventual reserves
  
  - 30m x 2m transects
  
  - ID, count, and estimate length of all observed fish species
  
  - Lengths converted to weights via allometric relationships 
  
  - Biomass density of each species for each transect estimated


---

# Data Coverage

```{r}
big_trends <- length_data %>% 
  left_join(site_data, by = c('site','side')) %>% 
  group_by(year) %>% 
  summarise(`Total Observations` = length(classcode),
            `Species Observed` = length(unique(classcode)),
            `Sites Surveyed` = length(unique(site)),
            `Number of Observers` = length(unique(observer)))

big_trends %>% 
  gather(variable, value, -year) %>% 
  ggplot(aes(year, value, color = variable)) + 
  geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2) +
  geom_line(show.legend = F) + 
  geom_point(show.legend = F) +
  facet_wrap(~variable, scales = 'free_y') + 
    scale_color_manual(values = wes_palette("Zissou")) + 
  scale_y_continuous(labels = comma)

```

---

# Spatial Coverage


```{r}

spatial_trend <- length_data %>% 
  left_join(site_data, by = c('site','side')) %>% 
  filter(region %in% c('ANA','SCI','SMI','SRI')) %>% 
  mutate(eventual_mpa = year_mpa > 0) %>% 
  group_by(year,site, side) %>% 
  summarise(n_obs = length(classcode),
            eventual_mpa = unique(eventual_mpa)) %>% 
  left_join(site_coords, by = c('site','side'))

ggmap::qmplot(longitude, latitude, color = eventual_mpa, data = spatial_trend) + 
  facet_wrap(~year) + 
  ggsci::scale_color_npg(name = 'Eventual MPA?') +
  theme_minimal() + 
  theme(axis.text = element_blank())

```

---

# Species Covered


```{r}

load(file = paste0(run_dir, '/length_to_density_data.Rdata'))

fished_species <- read_csv(file = file.path('../data','cfdw-channel-islands-catches.csv')) %>% group_by(classcode) %>% 
  summarise(catch = sum(pounds_caught, na.rm = T))


fished_species <- data_frame(classcode = unique(fished_species$classcode), fished = 1)


length_to_density_data <- length_to_density_data %>% 
    mutate(classcode = tolower(classcode)) %>% 
   left_join(life_history_data %>% select(classcode, commonname,targeted), by = 'classcode')  %>% 
  left_join(fished_species, by = c('classcode')) %>% 
  left_join(site_data, by = c('site','side')) %>%
  mutate(targeted = ifelse(fished == 1 & is.na(fished) == F & targeted == "Non-targeted", 'Targeted', targeted))

length_to_density_data %>% 
  group_by(classcode, commonname, targeted) %>% 
  summarise(n_obs = length(year)) %>% 
  ungroup() %>% 
  filter(is.na(commonname) == F, is.na(targeted) == F, str_detect(commonname,'YOY') == F) %>% 
    ungroup() %>% 
  mutate(commonname = fct_reorder(commonname, n_obs)) %>% 
  top_n(35, n_obs) %>% 
   ggplot(aes(commonname, n_obs, fill = targeted), color = 'black') +
  geom_col() + 
  coord_flip() + 
  theme(axis.text.y = element_text(size = 8))
```

---

# Density Data

```{r}


length_to_density_data %>% 
  ggplot(aes(log(mean_biomass_g + 1e-3))) + 
  geom_histogram() + 
  labs(x = 'Log density', y = 'Observations', caption = 'missing fish added in')

```

---

# Environmental Covariates


```{r}

conditions_data %>% 
  left_join(site_data %>% select(site, region), by = 'site') %>% 
  gather(variable, value, contains('mean_')) %>% 
  ggplot(aes(year,value, color = variable)) + 
  geom_violin(show.legend = F, aes(group = year, fill = variable), alpha = 0.5) + 
  geom_smooth(show.legend = F) +
  facet_grid(variable ~., scales = 'free_y')

```

---

# Environmental Covariates

```{r}

enso %>% 
  select(year, mean_enso) %>% 
  left_join(pdo %>% select(year, mean_pdo), by = 'year') %>% 
  filter(year >= 1999, year <=2013) %>% 
  gather(variable, value, -year) %>% 
  ggplot(aes(year, value, color = variable)) + 
  geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2) +
  geom_hline(aes(yintercept = 0)) +
  geom_line(show.legend = F) +
  geom_point(show.legend = F) + 
  facet_wrap(~variable)
  
```

---

# CFDW Catch Data

```{r}

ci_catches <- read_csv(file = file.path('../data','cfdw-channel-islands-catches.csv'))

ci_catches %>%
  filter(common_name != 'pacific sardine') %>%
  group_by(common_name, year) %>%
  summarise(catch = sum(pounds_caught))  %>%
  group_by(common_name) %>%
  mutate(total_catch = sum(catch)) %>%
  ungroup() %>%
  mutate(common_name = fct_reorder(common_name,total_catch )) %>%
  group_by(common_name) %>% 
  mutate(catch = (catch - mean(catch, na.rm = T)) / (2 * sd(catch, na.rm = T))) %>% 
  ungroup() %>% 
  filter(is.na(catch) == F) %>% 
  ggplot(aes(year,catch, color = common_name)) +
  geom_vline(aes(xintercept = 2003), color = 'red',
             linetype = 2) +
  geom_line( show.legend = F) + 
    facet_wrap(~common_name, as.table = F) + 
theme_minimal() +
  theme(axis.text = element_blank(), 
        panel.grid = element_blank())

```

---

# Key Data Issues

- Massive changes in sampling quantity and coverage

- Litte "pre" coverage

- A boatload of zeros

- Mother nature went nuts right along with MPAs

- How to include catch data 

---

class: inverse, center, middle
# Solutions

---

# Raw Trends


```{r}

length_to_density_data %>% 
  mutate(classcode = tolower(classcode)) %>% 
  filter(is.na(targeted) == F) %>% 
  group_by(year, targeted) %>% 
  summarise(mean_density = mean(mean_biomass_g, na.rm = T)) %>% 
  ggplot(aes(year, mean_density, color = targeted)) + 
  geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2) +
  geom_line() + 
  geom_point() + 
  ggsci::scale_color_npg() + 
  labs(title = 'Unfiltered mean densities')

```

---

# Raw Trends


```{r}
length_to_density_data %>% 
  filter(is.na(year_mpa) == F) %>% 
  mutate(classcode = tolower(classcode),
         eventual_mpa = year_mpa > 0) %>% 
  filter(is.na(targeted) == F) %>% 
  group_by(year, targeted,eventual_mpa) %>% 
  summarise(mean_density = mean(mean_biomass_g, na.rm = T)) %>% 
  ggplot(aes(year, mean_density, color = targeted)) + 
  geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2) +
  geom_line() + 
  geom_point() + 
  ggsci::scale_color_npg() + 
  facet_wrap(~eventual_mpa) +
  labs(title = 'Unfiltered mean densities')
```

---

# Data Filtering

- No 1999 (sampling protocols really changed)

- Only main channel islands (SMI, SRI, SCI, ANA)

- Species observed in at least 8 years, at least once pre MPA once post MPA

- No "young of the year" (recruits) - too hard to spot

---

# Filtered Trends

```{r}

abundance_models %>% 
  filter(data_source == 'length_to_density') %>% 
  mutate(commonname = fct_reorder(commonname, dim_data)) %>% 
  ggplot(aes(commonname, dim_data, fill = targeted)) + 
  geom_col(position = 'dodge') + 
  coord_flip() 
```

---

# Filtered Trends

```{r}
model_data <- abundance_models %>% 
  filter(data_source == 'length_to_density') %>% 
  select(classcode, data,commonname, targeted) %>% 
  unnest() %>% 
  mutate(year = factor_year %>% as.character() %>% as.numeric())

model_data %>% 
  filter(is.na(year_mpa) == F) %>% 
  mutate(classcode = tolower(classcode),
         eventual_mpa = year_mpa > 0) %>% 
  filter(is.na(targeted) == F) %>% 
  group_by(year, targeted,eventual_mpa) %>% 
  summarise(mean_density = mean(mean_biomass_g, na.rm = T)) %>% 
  ggplot(aes(year, mean_density, color = targeted)) + 
  geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2) +
  geom_line() + 
  geom_point() + 
  ggsci::scale_color_npg() + 
  facet_wrap(~eventual_mpa) +
  labs(title = 'Filtered mean densities')

```

---


class: inverse, middle, center
# Model-Palooza

---

# Strategy #1: Ignore the zeroes

- Unclear how to incorporate zeros into normal DiD estimator
  - Standard tobit/probit doesn't work in this case
  
- So, easier to ignore!

- Broader rationale: detection probability relatively independent of population size?
  - Not entirely unreasonable given focus on common species and challenges of survey technique
  
- Estimate DiD terms with a whole pile of covariates

---

# Strategy #1: Ignore the zeroes


```{r load-old-ahnold}

runfolder <- 'Working'

runpath <- paste('~/phd/ahnold/results/',runfolder,'/', sep = '')

load(paste(runpath,'regressions.Rdata', sep = ''))

load(paste(runpath,'plots.Rdata', sep = ''))

plot_list$mpa_effect_plot

```

---

# Strategy #1: Ignore the zeroes

```{r}
plot_list$in_mpa_effect_plot
```

---

# Strategy #1: Ignore the zeroes


```{r}

plot_list$mpa_anasci_effect_plot

```

---

# Strategy #1: Ignore the zeroes

- How to explain these results?

- Developed spatially explicit bio-economic MPA simulater (SPASM!)

- Assess densities with and without MPAs under wide range of life history and economics

- How likely, and under what circumstances, do we observe a net decrease in densities?

---


# Strategy #1: Ignore the zeroes
## What would explain results?

```{r}

sim_grid_plots$sim_plot[[1]]

```

---

# Strategy #1: Ignore the zeroes
## What would explain results?


```{r}

bad_mpa_foo <- function(x){
  
  year_mpa <- min(x$year[x$percent_mpa > 0])
  
  xp <- x %>% 
    filter(year >= year_mpa, year <= (year_mpa + 14)) %>% 
    select(year,experiment, ssb) %>% 
    spread(experiment, ssb) %>% 
    mutate(no_mpa_better = `no-mpa` > `with-mpa`)
  
  mpa_better = any(xp$no_mpa_better == T)
  
}

slice_grid <- slice_grid %>% 
  mutate(mpa_bad = map_lgl(mpa_experiment, bad_mpa_foo))

slice_grid %>% 
  ggplot(aes(mpa_bad)) + 
  geom_bar(position = 'stack') + 
  labs(title = 'MPA made things worse?')


```

---

# Strategy #1: Ignore the zeroes
## What would explain results?


```{r}

bad_mpa <- which(slice_grid$mpa_bad) 

sim_grid_plots[bad_mpa,]$sim_plot[[2]]

```

---

# Strategy #2: Zeros then DiD

- Rather than ignore zeros, incorporate into standard delta-GLM style "CPUE" standardization:
 
  - Control for non-abundance based reasons we either see zeros or see the densities we see

  - Observer, visibility, site, etc.
  
- Extract index of abundance controlling for above junk

- Fit DiD estimator for abundance estimates

  - Deals with zeros, species weighting, etc. 
  
  - Makes coefficients easier to understand
  
- Uncertainty a bit harder to deal with (MCMC fixes)

- Assumes indicies are right....

---

# Strategy #2: Zeros then DiD

```{r}

model_comparison_plots$comp_plot[[8]]

```

---

# Strategy #2: Zeros then DiD

```{r}

did_models$did_plot[[3]] + 
  labs(title = "D'oh!")

```

---


# What next?

Put on your reviewer hats....

Assuming these results hold...

- Something is happening but we can't really explain it

- MLPA has no detectable causal effect on fish densities

- Environmental signals swamp MPA effect

- Data just aren't able to answer this question...

- Classical MPA models do not reflect observations from Channel Islands

- Abandon ship

- Another way in?

---


