---
title: "Testing for Trophic Interactions using Convergent Cross Mapping"
author: "Owen Liu"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
```

## Introduction

Here we use convergent cross mapping (CCM) sensu Sugihara et al. 2012 *Science* to test for a significant causal signal between different trophic groups in monitoring data from the northern California Channel Islands. CCM is part of a larger suite of tools collectively known as empirical dynamic modeling, and many of those tools are implemented in the `rEDM` package.

```{r load packages}
require(tidyverse)
require(rEDM)
require(ggsci)

# load data
dat <- read_csv("C:/Users/oliu.MASONBEE/Documents/UCSB/Lab Meetings/trophic_group_densities.csv",col_types="cidddddd") %>%
  arrange(region,year)
```

## Process data for analysis

Time series have to be on the same scale and stationary. Split by region (island) and standardize

```{r standardize timeseries}
dat %>%
  select(-lag4_carnivore,-lag4_piscovore)%>%
  group_by(region)%>%
  mutate_at(vars(carnivore:planktivore),funs(norm=(.-mean(.))/sd(.)))->dat

dat %>%
  select(region,year,contains("norm"))%>%
  gather("group","density",-region,-year)%>%
  ggplot(aes(year,density,col=group))+
  geom_line()+
  scale_color_locuszoom(name="trophic group",labels=c("carnivore","herbivore","piscivore","planktivore"))+
  labs(x="year",y="normalized density")+
  facet_wrap(~region)

## split data by region to analye timeseries from each island separately
ana.dat <- dat %>% filter(region=="ANA")
sci.dat <- dat %>% filter(region=="SCI")
smi.dat <- dat %>% filter(region=="SMI")
sri.dat <- dat %>% filter(region=="SRI")

datnest <- dat %>% group_by(region) %>% nest()

```

## Embedding dimension

We have to test whether the variables can be properly embedded (i.e., if they have predictable manifold dynamics). We do this through a simplex forecasting test, using an individual timeseries' own lags to build a manifold. For each timeseries, the "best embedding dimension" is an approximation of the dimensionality of the dynamic system, in other words, the number of dimensions that define and predict the evolving states of the timeseries.

We assess best embedding dimension using `simplex` in `rEDM`, that tests multiple embedding dimensions and compares their ability to forecast the timeseries, using leave-one-out cross-validation.

```{r simplex embedding}
var_names <- c("carnivore_norm","herbivore_norm","piscivore_norm","planktivore_norm")

map(datnest$data,function(x) {
  simp.out <- map(var_names,function(y){
    temp <- x %>% unnest() %>% ungroup() %>% select(matches(y)) %>% as.data.frame()
    out <- simplex(as.numeric(temp[,1]),E=1:5,silent=T)
    out
  })
  simp.out
}) -> simplist

map(simplist,)
bind_rows(simplist) %>%
  ggplot(aes(E,rho))+
  geom_line(size=2)+
  facet_wrap(~spp,nrow=2,scales="free_y")+
  labs(x="Embedding Dimension (E)",y=expression(paste("Skill, ",rho)))+
  scale_x_continuous(breaks=seq(0,10,by=2))+
  guides(color=F)+
  theme(strip.text = element_text(size=8))
```

An inspection of the individual simplex model runs is not promising--these timeseries seem too short to have any evidence of predictability.

As another option, we could use all the data and stitch together separate samples using spatial convergent cross mapping (Clark et al. 2015 *Ecology*). This uses the data from the islands as "replicates", requiring the assumption that the islands are all part of the same "dynamic system", but allowing us to take advantage of the extra information to further resolve the reconstructure manifolds.

```{r simplex with regions combined}
datnorm <- dat %>% select(region,year,contains('norm')) %>% ungroup()

# have to record the segments corresponding to each "replicate" so simplex algorithm does not try to make predictions crossing time barriers
segs <- datnorm %>% mutate(ind=row_number()) %>% group_by(region) %>% summarise(first=first(ind),last=last(ind)) %>%
  select(-region)

regions.combined.simp.list <- map(var_names,function(x){
  temp <- datnorm %>% ungroup() %>% select(matches(x)) %>% as.data.frame()
  out <- simplex(as.numeric(temp[,1]),E=1:10,lib=as.matrix(segs),silent=T) %>%
    mutate(trophic=x)
  out
})

# plot the embedding dimension vs. rho, the Pearson's correlation between predictions and observations
bind_rows(regions.combined.simp.list) %>%
  ggplot(aes(E,rho,color=trophic))+
  geom_line(size=2)+
  facet_wrap(~trophic,nrow=2,scales="free_y")+
  geom_hline(yintercept = 0,color="black")+
  labs(x="Embedding Dimension (E)",y=expression(paste("Skill, ",rho)))+
  scale_x_continuous(breaks=seq(0,12,by=2))+
  scale_color_locuszoom()+
  guides(color=F)
```

With this stitched-together data, it still seems like perhaps only carnivore and herbivore dynamics are significantly predictable, given the data we have.

## Convergent Cross Mapping

With these two timeseries, we can test for unidirectional or bidirectional causality using convergent cross mapping. In the simplex forecasting above, we used multiple lags of a single timeseries (E lags) to make predictions.

So far, we have established the predictability of single variables. In this step, we relate the variables to each other, asking whether there is evidence of causal linkages. We do not assume causal interactions--we test for them.

Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence. In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series. Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping"). CCM works just like univariate simplex projection that we did in Step 1, except that separate variables are used for library and prediction vectors. If the attractor built from the first variable can accurately predict the dynamics of the second variable, we can claim that the second variable has a causal influence on the first.  In simple terms, the *causal effect of A on B is determined by how well B cross-maps A*. In this way, the inference from cross-mapping is the converse direction of causation. In this example, if herbivores drive carnivores,the dynamic information from the herbivore time series should be reflected in the carnivore dynamics, and carnivores should significantly cross-map the herbivores.

There are two criteria for CCM to establish causality: 

* First, and most obviously, predictive cross-map skill using all available data should be significantly greater than zero. 
* Second, that predictability should be convergent.  Convergence means that cross-mapped estimates improve with library length (the number of state-space vectors used to build the attractor), because the attractor is more fully resolved and therefore estimation error should decline. Convergence is key to distinguishing causation from simple or spurious correlation. If two variables are spuriously correlated and not causally linked, CCM should fail to satisfy this second criterion.

The CCM algorithm uses a random sampling method to test multiple "versions" of each library size, sampling a subset from the supplied library vectors to establish confidence intervals around prediction skill. We again use leave-one-out cross-validation to prevent in-sample fitting.

```{r ccm carn herb}
# cross map carnivore to herbivore
# inspect the output of simplex from the previous step and use the best embedding dimension (highest rho) for the herbivore time series

tempE <- 8

temp <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=tempE,lib_column= 'carnivore_norm',target_column = 'herbivore_norm',lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)

c_xmap_h <- temp %>%
  group_by(lib_size)%>%
  summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
  ungroup()%>%
  ggplot(aes(lib_size,rhomean))+
  geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
  geom_line(color="darkorchid3")+
  labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title="carnivore xmap herbivore")

c_xmap_h

# cross map herbivore to carnivore
# inspect the output of simplex from the previous step and use the best embedding dimension (highest rho) for the carnivore time series

tempE <- 6

temp <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=tempE,lib_column= 'herbivore_norm',target_column = 'carnivore_norm',lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)

h_xmap_c <- temp %>%
  group_by(lib_size)%>%
  summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
  ungroup()%>%
  ggplot(aes(lib_size,rhomean))+
  geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
  geom_line(color="darkorchid3")+
  labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title="herbivore xmap carnivore")

h_xmap_c
```

From this initial analysis, there is no evidence that carnivores drive herbivores, and a minor amount of evidence that herbivores drive carnivores.